Bootstrapping Domain Specific Languages for Neurally-Guided Bayesian Program Learning

### 2018-02-08 paper.tex
updates for list domain

\subsection{List functions}
Our list function domain consists of hand-designed concepts from which we
sample tasks. All tasks are solved by functions that take as input an
integer or a list of integers, and have as output either a Boolean, an
integer, a list of Booleans, or a list of integers.  Examples of these
tasks are in Table~\ref{listexamples}.

@@

% TODO: table of list results, show some induced primitives

@@

For each domain,
we are interested both in how many tasks the
agent can solve and how quickly it can find those solutions.
For symbolic regression, we also care about the quality
of the solution, as measured by the likelihood model $\probability[x|p]$.
Concretely, this cashes out as measuring how many
real numbers were used in the programs: e.g. did the model correctly explain a linear function
using two numbers, or did it introduce extraneous parameters?
Table~\ref{baselineComparisons}
compares our model against these baselines.
We see that our full model consistently
beats the baselines,
and that the recognition model 
consistently increases the number of solved held-out tasks.
Lesioning the recognition model also slows down the convergence of the algorithm,
taking more wake/sleep cycles to reach a given number of tasks solved (Figure~\ref{learningCurves}).
This data supports our view of the recognition model as a way of accelerating the search over programs.
\begin{table}
%% This radically shrinks the table
\tabcolsep=4pt
\renewcommand{\arraystretch}{0.5}
\begin{tabular}{llllll}
  \toprule& Ours&Ours--NN  & RF/DC & PCFG & Enum

  \\\midrule\multicolumn{6}{c}{\emph{Boolean Circuits}}\\\midrule
  \% solved &
  84\% &76\%& 63\%&62\%&62\%\\
  Solve time&  $0.9$s&  1.1&1.2&1.3&1.3

  \\\midrule\multicolumn{6}{c}{\emph{Symbolic Regression}}\\\midrule
  \% solved&    98\% &94\%&0.7\%&0\%&2\% \\
  Solve time&  $2.7$s& $2.8$s  &$3.6$s&--&$2.1$s\\
  MDL (nats)&   $9.3$ &$11.0$&$2.3$&--&$4.6$

  \\\midrule\multicolumn{6}{c}{\emph{String Editing}}\\\midrule
  \% solved&99\% &82\% &9\%&24\%&18\%\\
  Solve time&  0.9s&2.2s&2.7s&2.3s&2.4s
  \\\bottomrule
  \end{tabular}
\caption{\% solved: fraction solved with 5 sec timeout. Solve time: averaged over solved tasks. MDL: $-\expect\left[\probability[x|p] \right]$. For domains other than symbolic regression MDL is 0 nats. For symbolic regression MDL is a proxy for \# of continuous degrees of freedom.}\label{baselineComparisons}  \end{table}

>>>>

\subsection{List functions}
Our list function domain consists of hand-designed concepts from which we
sample tasks, so there is no ``ground truth'' formal DSL from which these
tasks are generated. Additionally, few of these tasks are concatenative ---
many of them represent distinct operations. We found that this domain was
difficult to tackle without starting from a relatively rich DSL, and even
then it was still hard to solve all tasks. All tasks are solved by functions
that take as input an integer or a list of integers, and have as output
either a Boolean, an integer, a list of Booleans, or a list of integers.
Examples of these tasks are in Table~\ref{listexamples}.

@@

\begin{table}[t]
  \centering
  Subset of learned DSL\\
  \begin{tabular}{l}
    \toprule
    $(f_0$ \code{i $\ell$}$) = $\code{(singleton (index i $\ell$))}\\
    \hspace{0.5cm}($f_0$: \emph{Put the $i$-th index into a list})\\
    $(f_1$ \code{i $\ell$}$) = $\code{(++ $\ell$ ($f_0$ i $\ell$))}\\
    \hspace{0.5cm}($f_1$: \emph{Append the $i$-th index})\\
    $(f_2$ \code{n $\ell$}$) = $\code{(any (lambda (x) (= n x)) $\ell$)}\\
    \hspace{0.5cm}($f_1$: \emph{Whether $n$ appears in the list})\\
    $(f_3$ \code{i $\ell$}$) = $\code{(index (negate i) (sort $\ell$))}\\
    \hspace{0.5cm}($f_3$: \emph{Get the $i$-th largest number})\\
    $(f_4$ \code{n $\ell$}$) = $\code{(mapi (lambda (i x) (mod x n)) $\ell$)}\\
    $(f_5$ \code{k $\ell$}$) = $\code{(mapi (lambda (i x) (+ x k)) $\ell$)}\\
    $(f_6$ \code{k n $\ell$}$) = $\code{($f_4$ n ($f_5$ k $\ell$)}\\
    \hspace{0.5cm}($f_6$: \emph{Caesar shift of $k$ in integers modulo $n$})\\
  \bottomrule
  \end{tabular}
  \caption{Some induced primitives from the list DSL}\label{listinduced}
\end{table}

@@

For each domain,
we are interested both in how many tasks the
agent can solve and how quickly it can find those solutions.
For symbolic regression, we also care about the quality
of the solution, as measured by the likelihood model $\probability[x|p]$.
Concretely, this cashes out as measuring how many
real numbers were used in the programs: e.g. did the model correctly explain a linear function
using two numbers, or did it introduce extraneous parameters?
Table~\ref{baselineComparisons}
compares our model against these baselines.
We see that our full model consistently
beats the baselines,
and that the recognition model 
consistently increases the number of solved held-out tasks.
Lesioning the recognition model also slows down the convergence of the algorithm,
taking more wake/sleep cycles to reach a given number of tasks solved (Figure~\ref{learningCurves}).
This data supports our view of the recognition model as a way of accelerating the search over programs.

\begin{table}
%% This radically shrinks the table
\tabcolsep=4pt
\renewcommand{\arraystretch}{0.5}
\begin{tabular}{llllll}
  \toprule& Ours&Ours--NN  & RF/DC & PCFG & Enum

  \\\midrule\\\multicolumn{6}{c}{\emph{Boolean Circuits}}\\\midrule
  \% solved &
  84\% &76\%& 63\%&62\%&62\%\\
  Solve time&  0.9s&  1.1s&1.2s&1.3s&1.3s

  \\\midrule\\\multicolumn{6}{c}{\emph{Symbolic Regression}}\\\midrule
  \% solved&    98\% &94\%&0.7\%&0\%&2\% \\
  Solve time&  2.7s& 2.8s  &3.6s&--&2.1s\\
  MDL (nats)&   9.3 &11.0&2.3&--&4.6

  \\\midrule\\\multicolumn{6}{c}{\emph{String Editing}}\\\midrule
  \% solved&99\% &82\% &9\%&24\%&18\%\\
  Solve time&  0.9s&2.2s&2.7s&2.3s&2.4s

  \\\midrule\\\multicolumn{6}{c}{\emph{List functions (base DSL)}}\\\midrule
  \% solved&52\% &52\% &40\%&44\%&42\%\\
  Solve time&  0.4s&0.7s&1.2s&0.9s&1.0s

  \\\midrule\\\multicolumn{6}{c}{\emph{List functions (rich DSL)}}\\\midrule
  \% solved&86\% &84\% &60\%&74\%&70\%\\
  Solve time&  0.8s&0.7s&1.0s&1.0s&1.1s
  \\\bottomrule
  \end{tabular}
\caption{\% solved: fraction solved with 5 sec timeout. Solve time: averaged over solved tasks. MDL: $-\expect\left[\probability[x|p] \right]$. For domains other than symbolic regression MDL is 0 nats. For symbolic regression MDL is a proxy for \# of continuous degrees of freedom.}\label{baselineComparisons}  \end{table}
